---
title: "Project-Heart Disease Prediction"
author: "Chaoying Luo, Yiyang Sheng, Ying Liu, Ruitong Liu, Bo Yang"
date: "2021/4/29"
output: html_document
---


```{r}
library(tidyverse) 
library(ggplot2) 
library(caret) # Machine learning
library(caTools)
library(gridExtra)   
library(grid)        
library(pROC)
```


```{r}
data = read.csv("test.csv")
data = data[,-1]
head(data,6)

# correct the data type
data$Sex_male = as.factor(data$Sex_male)
data$currentSmoker = as.factor(data$currentSmoker)
data$BPMeds = as.factor(data$BPMeds)
data$prevalentStroke = as.factor(data$prevalentStroke)
data$prevalentHyp = as.factor(data$prevalentHyp)
data$diabetes = as.factor(data$diabetes)
data$TenYearCHD = as.factor(data$TenYearCHD)
str(data)
```


```{r}
# split the data into training set and test set
set.seed(123)
split_data = sample.split(data$TenYearCHD, SplitRatio = 0.7)
training_data = data[split_data==T,]
test_data = data[split_data==F,]
nrow(training_data)
nrow(test_data)
```

```{r}
# run logitic regression
logistic_model = glm(TenYearCHD~., data = training_data, family = "binomial")
summary(logistic_model)
```

Significant variables include Sex_male, age, cigsPerDay, prevalentStroke, sysBP, and glucose. We can run logitic regression again with these significant variables.

```{r}
# run logitic regression again with significant variables
logistic_model_new = glm(TenYearCHD~Sex_male+age+cigsPerDay+prevalentStroke+sysBP+glucose, data = training_data, family = "binomial")
summary(logistic_model_new)
```

```{r}
# predict TenYearCHD on test data
pred_model =  predict(logistic_model,newdata = test_data[,-15],type = "response")
summary(pred_model)

pred_test_data = ifelse(pred_model>0.5,1,0)
```

```{r}
cm = confusionMatrix(pred_test_data, test_data$TenYearCHD)
cm$table

# extract the confusion matrix values as data.frame
cm_d = as.data.frame(cm$table)
# confusion matrix statistics as data.frame
cm_st = data.frame(cm$overall)
# round the values
cm_st$cm.overall = round(cm_st$cm.overall,2)

# here we also have the rounded percentage values
cm_p = as.data.frame(prop.table(cm$table))
cm_d$Perc = round(cm_p$Freq*100,2)

# plotting the matrix
cm_d_p =  ggplot(data = cm_d, aes(x = Prediction , y =  Reference, fill = Freq))+
  geom_tile() +
  geom_text(aes(label = paste("",Freq,",",Perc,"%")), color = 'red', size = 5) +
  theme_light() +
  guides(fill=FALSE) 

# plotting the stats
cm_st_p =  tableGrob(cm_st)

# all together
grid.arrange(cm_d_p, cm_st_p,nrow = 1, ncol = 2, 
             top=textGrob("Confusion Matrix and Statistics",gp=gpar(fontsize=20,font=1)))
# evaluation
confusion_matrix = table(pred_test_data,test_data$TenYearCHD)
acc = (sum(diag(confusion_matrix))/sum(confusion_matrix))*100
paste("Overall accuracy with threshold 0.5 is",acc,"%")

# roc
pROC_obj = roc(test_data$TenYearCHD,pred_model, 
               smoothed = TRUE,
               # arguments for ci
               ci=TRUE, ci.alpha=0.9, stratified=FALSE,
               # arguments for plot
               plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
               print.auc=TRUE, show.thres=TRUE)
```

Though the accuracy is 84.97%, this model is not predicting people who have the risk of heart disease correctly.
As we care more about people who have chances of heart disease, we can try a different threshold and compute accuracy only for those who have 10 year risk of coronary heart disease.

```{r}
# threshold = 0.1
pred_test_data_lowthreshold = factor(ifelse(pred_model>0.1,1,0))

cm = confusionMatrix(pred_test_data_lowthreshold, test_data$TenYearCHD)
cm$table

# extract the confusion matrix values as data.frame
cm_d = as.data.frame(cm$table)
# confusion matrix statistics as data.frame
cm_st = data.frame(cm$overall)
# round the values
cm_st$cm.overall = round(cm_st$cm.overall,2)

# here we also have the rounded percentage values
cm_p = as.data.frame(prop.table(cm$table))
cm_d$Perc = round(cm_p$Freq*100,2)

# plotting the matrix
cm_d_p =  ggplot(data = cm_d, aes(x = Prediction , y =  Reference, fill = Freq))+
  geom_tile() +
  geom_text(aes(label = paste("",Freq,",",Perc,"%")), color = 'red', size = 5) +
  theme_light() +
  guides(fill=FALSE) 

# plotting the stats
cm_st_p =  tableGrob(cm_st)

# all together
grid.arrange(cm_d_p, cm_st_p,nrow = 1, ncol = 2, 
             top=textGrob("Confusion Matrix and Statistics",gp=gpar(fontsize=20,font=1)))

acc_for_1 = (159/(159+34))*100
paste("Accuracy for those who have 10 year risk of coronary heart disease with threshold 0.1 is",acc_for_1,"%")
```

Now as we have a good accuracy for predicting people with chances of heart disease in next 10 years, we can treat them and guide them well in advance.
